---
title: "Chapter 20, problems 31 and 42"
output: html_document
runtime: shiny
---

```{r, include = FALSE}
#Including packages
require(magrittr)
require(ggplot2)
require(gridExtra)
## we want to use the read all funcion here
readFile <- function(fileLocation, dlm = "\t"){
  testData <- readLines(fileLocation)
  
 testData <- gsub("\\!|\\$|\\(|\\)|\\[|\\]|\\{|\\}|\\#|\\'|\\%|\\~|\\?|\\@|\\^|\\*|\\&|\\,",
                 "",
                 testData)
 
  testData <- gsub(" ", "", testData)
  
  framedData <- read.table(textConnection(testData), header = TRUE, sep = dlm)
  
  return(framedData)
}

location <- "C:\\Users\\Jonathan\\Google Drive\\Stats Camp\\Stine&Foster\\Data by Chapter\\Chapter 20\\20_walmart\\20_walmart.txt"

walmart <- readFile(location)

```
### Problem 31, Part A

```{r}

ggplot(data = walmart, aes(x = Date, y = NetIncome)) +
  geom_point(color = "red") +
  geom_line(color = "blue")

```

The pattern isn't linear. The points gradually spread out as time goes on. It makes a megaphonic patern.

### Problem 31, Part B: linear fit

```{r}

ggplot(data = walmart, aes(x = Date, y = NetIncome)) +
  geom_point() +
  geom_smooth(method = "lm")

walmartLM <- lm(NetIncome ~ Date, data = walmart)

summaryWalmart <- walmartLM %>% summary

```

The linear fit is: $NetIncome = `r round(summaryWalmart$coefficients[1,2], digits = 4)` * Date + `r round(summaryWalmart$coefficients[1,1], digits = 4)`$  


### Problem 31, Part C: Residual Plot

```{r}
walmartResid <- resid(walmartLM)
walmart$resid <- walmartResid

ggplot(data = walmart, aes(x = Date, y = resid)) +
  geom_point(color = "red") 
```

There is a patern in the residuals. It has is a megaphonic patern to it which indicates a non-linear patern.

### Problem 31, Part D: Comparison on the log and normal Scales.

```{r}
normalPlot <- ggplot(data = walmart, aes(x = Date, y = NetIncome)) +
                    geom_point() +
                    geom_smooth(method = "lm")

logPlot <- ggplot(data = walmart, aes(x = Date, y = LogOpIncome)) +
                  geom_point() +
                  geom_smooth(method = "lm")

grid.arrange(normalPlot, logPlot)
```

Looking at the two graphs together. The Log Income shows a mostly uniform residual size from the main line. The residuals on the Net income graph show more distance as the date increase. The better option is the log transformed data.

### Problem 31, Part E the fit for log equation

```{r}

WalmartLogLM <- lm(LogOpIncome ~ Date, data = walmart)

summaryLogLM <- WalmartLogLM %>% summary

```

The equation for the new line: $Date = `r summaryLogLM$coefficients[2,1]` * Date + `r summaryLogLM$coefficients[1,1]`$

The slope `r summaryLogLM$coefficients[2,1]` means that for increase of one day we expect the LogIncome to increase by `r summaryLogLM$coefficients[2,1]`.

### Problem 31, Part F residual patterns

```{r}
logResid <- resid(summaryLogLM)

walmart$logResid <- logResid

ggplot(data = walmart, aes(x = Date, y = logResid)) +
  geom_point(color = "blue")

```

The fourth quarter is always higher than the the other quarters. It's a little less bowed than the non transformed model.

### Problem 31, Part G

This The better choice is the log transformed data. The residual plot is better and it also successfully explains more of the data than the other model. 

### Problem 42 

#### Motivation
##### Part A

```{r, include = FALSE}

filePath <- "C:\\Users\\Jonathan\\Documents\\GitHub\\AnalyticsData\\Stine and Foster\\Data by Chapter\\Chapter 20\\20_4m_philadelphia\\20_4m_philadelphia.txt"

philly <- readFile(filePath)
```

If they can show a negitive correlation between crime rates and housing prices they'd be able to justify recruiting more police in an attept to raise property values.

##### Part B

The equation wouldn't be causal because "Correlation isn't Causation". A good equation indicates that crime rate can be used to predict housing prices but doesn't cause them. The property value of a shanty in the middle of a desert with a crime rate of zero doesn't mean that the property is valuable. 

#### Method
##### Part C

The community leader would use crime as the explanitory varible. His or her end goal is to try and incease spending on the police. Useing the crime rate to explain the housing prices makes the most convincing argument. Using property value to explain crime would be more likely to pull funding to help citizens build nicer homes or raise property values.


##### Part D

```{r}
philly <- philly[philly$Name != "PhilaCC",]
```
Economic Data tends to be non-linear so we don't expect a linear relationship in this case. 

#### Mechanics
##### Part E
```{r}

ggplot(data = philly, aes(x = CrimeRate, y = HousePrice)) +
  geom_point(color = "blue")

```

The relationship is strong-negitive and non-linear

##### Part F
```{r}
phillyLM <- lm(HousePrice ~ CrimeRate, data = philly)

phillyLMSummary <- summary(phillyLM)
```

The slope is: `r  phillyLMSummary$coefficients[2,1]`

The Intercept is: `r phillyLMSummary$coefficients[1,1]`

The R^2^ value is: `r phillyLMSummary$r.squared`

`r phillyLMSummary$r.squared %>% round(digits = 4)* 100` percent of the variation in the housing price can be explained Crime Rate.

The Standard Error: `r phillyLMSummary$coefficients[2, 2]`

##### Part G
```{r}
ggplot(data = philly, aes(x = X1000.Crime, y = CrimeRate)) +
  geom_point(color = "red")

```

#### Message
##### Part H

The Summary using the reciprocal (1/x) 

```{r}
phillyTransLMSummary <- lm(X1000.Crime ~ CrimeRate, data = philly) %>% summary

phillyTransLM <- lm(X1000.Crime ~ CrimeRate, data = philly)
```

**Non-Transformed**

*Slope:* `r  phillyLMSummary$coefficients[2,1]`

*Intercept:* `r phillyLMSummary$coefficients[1,1]`

*R^2^:* `r phillyLMSummary$r.squared`

*Standard Error:* `r phillyLMSummary$coefficients[2, 2]`


**Transformed**

*Slope:* `r  phillyTransLMSummary$coefficients[2,1]`

*Intercept:* `r phillyTransLMSummary$coefficients[1,1]`

*R^2^:* `r phillyTransLMSummary$r.squared`

*Standard Error:* `r phillyTransLMSummary$coefficients[2, 2]`

```{r}
TransResids <- resid(phillyTransLM)

NonTransResids <- resid(phillyLM)

```



##### Part I

$Housing Price = `r  phillyTransLMSummary$coefficients[2,1]` * CrimeRate + `r phillyTransLMSummary$coefficients[1,1]`$

##### Part J

```{r}
inputPanel(
  numericInput("one", "First Value to be predicted", value = 1),
  numericInput("two", "Second value to be predicted", value = 2)
)
inputPanel(
  numericInput("three", "First Value to be predicted", value = 11),
  numericInput("four", "Second value to be predicted", value = 12)
)

renderPlot({
  ggplot(data = philly, aes(x = X1000.Crime, y = CrimeRate)) +
  geom_point(color = "red") + 
  geom_vline(xintercept = input$one) +
  geom_vline(xintercept = input$two) +
  geom_vline(xintercept = input$three) +
  geom_vline(xintercept = input$four) 
})

renderText({
  
  crimePred <- data.frame(CrimeRate = c(input$one, input$two, input$three, input$four))
  
  predictions <- predict.lm(phillyTransLM, newdata = crimePred)
  
  differences <- c(predictions[1] - predictions[2], predictions[3] - predictions[4])
  
  paste("The difference between the first two is: ", differences[1],
        "The difference between the second two is: ", differences[2])
})

```

Using the linear model to predict the differences we get the same difference between the points. 






