---
title: "Chapter 19 Problem 38"
author: "Jonathan Nelson"
date: "July 22, 2015"
output: html_document
runtime: shiny
---

### Part A: Scatter Plot, Sales v. Volume (gallons)

```{r, include = FALSE}
require(shiny)
require(ggplot2)
require(reshape2)
require(magrittr)
require(devtools)

fileName <- "19_convenience"

gasURL <- paste0("http://marriottschool.net/teacher/govfinance/", 
                 "resources/dataSets/stineFoster/Text/",
                 fileName, ".txt")

 gas <- read.table(gasURL,
             header = TRUE,
             sep = "\t"
             )
 
names(gas) <- c("sales", "gallons", "washes", "day", "cents", "weekend")

gas$weekend <- factor(gas$weekend, labels = c("WeekDay", "WeekEnd"))
```


```{r}
ggplot(data = gas, aes(x = gallons, y = sales, color = weekend)) +
        geom_point() +
        labs(x = "Gallons", y = "sales") +
        guides(color = guide_legend(title = "Day Type")) +
        geom_hline(yintercept = mean(gas$sales)) +
        geom_vline(xintercept = mean(gas$gallons))
```

### Part B: Estimate the linear equation, using the least-Square method

```{r}
gasLM <- lm(sales ~ gallons, data = gas)
```

The estimated equation is:
$sales = `r round(gasLM$coefficients[1],0)` + `r round(gasLM$coefficients[2], 2)` * gallons$

### Part C: Interpret the summary values r^2^ 

```{r}
modelSummary <- gasLM %>% summary

```

**The r-squared value is:** `r round(modelSummary$r.squared, 2)`

`r round(modelSummary$r.squared*100, 2) ` percent of the variation in sales can be explained gallons sold on a given day.

**The standard error of the regression:** `r round(modelSummary$coefficients[2,2], 2)`

`r round(modelSummary$coefficients[2,2], 2) ` is the average measurement error from the regression line. 

### Part D: Different predictions

```{r}

inputPanel(
  numericInput("one", "First Value to be predicted", value = 3500),
  numericInput("two", "Second value to be predicted", value = 4000)
)

renderText({
  gasGallons <- data.frame(gallons = c(input$one, input$two))
  
  predictions <- predict.lm(gasLM, newdata = gasGallons)
  
  paste("The Difference between the two is: ",
       round(predictions[2] - predictions[1], 2),
       ".", sep = "")
})

```

### Part E: Alternative units, change gallons to liters and dollars to Canadian dollars

```{r, include = FALSE}

gas$liters <- gas$gallons * 3.7854
gas$cDollars <- gas$sales * 1.1

litersLm <- lm(cDollars ~ liters, data = gas)

litersLm %>% summary

```

**The new estimated equation is:** $Sales = `r round(litersLm$coefficients[1],2)` +
`r round(litersLm$coefficients[2], 2)` * Liters$.

### Part F: Correlation and causation

A key component of this problem is the word 'causation'. "Correlation isn't causation" is an oft quoted statisticians maxim.` Causation is a word that we avoid using in regression/statistics in general. The best way to phrase our discovery is to put things interms of our response and explanatory variables.

We'd state that our response variable can be explained by our explanatory variable. So the amount sold can be explained by the number of gallons sold. How much explaining our explainitory variables is the R^2^ value. 

### Part G: Predicted values

```{r, include = FALSE}
newData <- data.frame(gallons = 4165)

pred4165 <- predict(gasLM, newdata = newData)

diffDat <- 1774 - pred4165
```

Our sales for the day were `r abs(round(diffDat, 2))` less than we expected.

### Part H: Residuals plot

```{r}
gas$residuals <- resid(gasLM)

ggplot(data = gas, aes(x = gallons, y = residuals, color = weekend)) +
  geom_point()

```

**The residual mean:** `r round(mean(resid(gasLM)), 2)`

**The residual sd:** `r round(sd(resid(gasLM)), 2)` 







